---
title: "Exercises"
output: html_notebook
---

1.What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?

```{r}
ggplot(data = mpg) +
  geom_line(mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(x = displ, y = hwy))

```

```{r}
ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(x = displ, y = hwy))

```

```{r}
ggplot(data = mpg) +
  geom_area(mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg) +
  geom_histogram(mapping = aes(x = displ))
```

2.Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

3.What does show.legend = FALSE do? What happens if you remove it?
Why do you think I used it earlier in the chapter?

It eliminates the variable legend around the graph. If we remove it the the variable legend will occur.

4. What does the se argument to geom_smooth() do?

It either eliminates or represents the deviation around the chart.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

```

```{r}
ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

No, because the layers inherit the configuration from ggplot.

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = drv, size = 5)) + geom_smooth(mapping = aes(x = displ, y = hwy, group = drv, size = 5, se = FALSE))
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(size = 4) + 
  geom_smooth(size = 4, se = FALSE )
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = drv, size = 4)) + 
  geom_smooth(size = 4, se = FALSE )
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, group = drv)) + 
  geom_point(size = 4) + 
  geom_smooth(size = 4, se = FALSE )
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = drv), size = 4) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv), size = 2, se = FALSE )
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "white", size = 4) + geom_point(mapping = aes(x = displ, y = hwy, color = drv), size = 3)
```

What is the default geom associated with stat_summary()? How could you rewrite the previous plot to use that geom function instead of the stat function?

geom_pointrange, 

```{r}
ggplot(data = diamonds) +
  geom_pointrange(mapping = aes(x = cut, y = depth),
                  stat = "summary",
                  fun.ymin = min,
                  fun.ymax = max,
                  fun.y = median)
```

`geom_col` leaves the data as it is. `geom_bar()` creates two variables (count and prop) and then graphs the count data on the y axis.

```{r}
?stat_smooth
```

Most geoms and stats come in pairs that are almost always used in concert. Read through the documentation and make a list of all the pairs. What do they have in common?
```{r}

```

	
geom	stat
area	identity
ribbon	identity
bar	count
col	identity
bin2d	bind2d
blank	identity
boxplot	boxplot
contour	contour
count	sum
crossbar	identity
errorbar	identity
errorbarh	identity
linerange	identity
pointrange	identity
curve	identity
segment	identity
density	density
density_2d	density2d
freqpoly	bin
histogram	bin
hex	binhex
jitter	identity
label	identity
path	identity
line	identity
map	identity
point	identity
polygon	identity
quantile	quantile
raster	identity
rect	identity
rug	identity
smooth	smooth
spoke	spoke
step	identity
text	identity
tile	identity
violin	ydensity

```{r}
stat_smooth()
```
What variables does stat_smooth() compute? What parameters control its behaviour?

`stat_smooth()` computes the y, the predicted value of y for each x value.

```{r}
ggplot(data = diamonds, aes(x = price, y = carat)) + stat_smooth()
```


the se of that value predicted, together with the upper and lower bound of that point prediction.

In our proportion bar chart, we need to set group = 1. Why? In other words what is the problem with these two graphs?

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))
```


```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..count.. / sum(..count..), fill = color))
```

**














