---
title: "Exercises"
output: html_notebook
---

1.What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?

```{r}
ggplot(data = mpg) +
  geom_line(mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(x = displ, y = hwy))

```

```{r}
ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(x = displ, y = hwy))

```

```{r}
ggplot(data = mpg) +
  geom_area(mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg) +
  geom_histogram(mapping = aes(x = displ))
```

2.Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

3.What does show.legend = FALSE do? What happens if you remove it?
Why do you think I used it earlier in the chapter?

It eliminates the variable legend around the graph. If we remove it the the variable legend will occur.

4. What does the se argument to geom_smooth() do?

It either eliminates or represents the deviation around the chart.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

```

```{r}
ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

No, because the layers inherit the configuration from ggplot.

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = drv, size = 5)) + geom_smooth(mapping = aes(x = displ, y = hwy, group = drv, size = 5, se = FALSE))
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(size = 4) + 
  geom_smooth(size = 4, se = FALSE )
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = drv, size = 4)) + 
  geom_smooth(size = 4, se = FALSE )
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, group = drv)) + 
  geom_point(size = 4) + 
  geom_smooth(size = 4, se = FALSE )
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = drv), size = 4) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv), size = 2, se = FALSE )
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "white", size = 4) + geom_point(mapping = aes(x = displ, y = hwy, color = drv), size = 3)
```















